\chapter{Conclusion and Final Thoughts}
\label{sec:conclusion}
Sigfox's uplink and downlink have been successfully reconstructed to a point where all relevant communication between object and base station is understood.
Both Sigfox object and base station can be emulated.
With \texttt{librenard} and \texttt{renard}, a reference implementation of an alternative Sigfox network stack has been created and its functionality has been tested.
Physical layer modulation and demodulation has been realized prototypically using an \gls{sdr} and python scripts.
None of the aspects of the Sigfox protocol are novel per se, Sigfox's innovation lies in the combination of known technologies into a wireless communication standard and the business aspects of marketing an \gls{iot} network.

All of Sigfox's claims regarding security-related aspects that were investigated were confirmed in a preliminary security analysis.
It has been found that neither uplink nor downlink payloads are confidential, but that was also never promised by Sigfox.
All in all, these results seem to be a good sign for the state of Sigfox security and can be seen as affirmation of the security-wise superiority of \gls{lpwan} networks in general and Sigfox in particular over traditional technologies like WiFi.
Certainly, improvements upon the current state of affairs regarding confidentiality are possible.
For example, the proposals in \cite{otpsecurity} include the use of \glspl{otp} for secure end-to-end encryption.

Topics for further work might be the yet uninvestigated \gls{oob} uplink frame and the implementation of a Sigfox object based on dedicated, commercially available \gls{rf} chips for \gls{unb} communications such as the STMicroelectronics S2-LP or the Texas Instruments CC1125.
An alternative implementation of a base station that is compatible with the Sigfox protocol might be also be of interest.
If these two components were present, an open sourced implementation of an \gls{lpwan} network that operates in parallel to Sigfox, but that is based on the Sigfox technology could be realized.
